Lens Modeling in \sw

### 2014-06-15 text/partA.tex
updated plot generation scripts input: kernel to inclease size of magenta parts mod: cut white border sim: plot without white border
editing

Galaxies and clusters of galaxies are considered to be the largest structures in the known Universe. When one questions their formation,  the present picture suggests that gravitational instabilities  (observable as fluctuations of the microwave background) %TODO \todo[red]{not too sure of the relevance of this statement} 
 - caused a collapse into gravitationally bound structures. %TODO\todo[yellow]{clarification needed: are they becoming gravitationally bound structures, or are they collapsing into objects that are already gravitationally bound? I know it is most probably the later, but it is a bit ambiguous}   
The initial collapse would have been dominated by dark matter, but once deep wells had appeared in the gravitational potential, gas would have started falling into those wells. 
%TODO \todo[red]{not too sure of this wording}
Through atomic processes,
%TODO \todo[yellow]{perhaps we need to clarify those processes or are they generally know?}
 the gravitational potential energy of this gas would of been radiated away, leading to the formation of the largest stars. 
%TODO \todo[red]{or perhaps just stars?}
In turn, galaxies and clusters of galaxies would form, all the time merging and changing their form.
This rough outline is now well established, but in order to satisfy our understanding of this process, there are many puzzles  that remain to be solved. For progress to be made, we need to extend our knowledge of dark matter. Gravitational lensing has proven as a particularly attractive way of studying dark matter as it is only dependant on mass.
%TODO \todo[yellow]{silly question, but shouldn't we briefly explain why it is only dependant on mass? or should everyone already reading this paper know that? or is it explained later?}

There are many manifestations of gravitational lensing in astrophysics,
but the most spectacular are multiple images.  Gravitational lensing
depends on the sky-projection mass density, and this fact has a
counter-intuitive consequence: nearby galaxies do not produce multiple
images, because their projected density is not high enough, but beyond
100~Mpc or so, most galaxies are potential lenses.  Even so, for a
potential lens to really act as a lens, a background source in the close-by
region is required.  So, in practice galaxies are very rarely observed
as lens.  The same tendency applies to clusters of galaxies, but since
clusters are bigger, they are more likely lens-like.

In lensing galaxies, the observable features usually involve a single
object multiply imaged.  It is therefore desirable to have large
numbers of lenses, and this is the strategy researchers are following.  The
largest single study so far uses 58 lenses \citep{2009ApJ...703L..51K}.
Other studies have compared lens models in more detail with stellar
mass \citep{2011ApJ...740...97L} or used time-delay information from
lensed quasars to infer the Hubble time
\citep{2008ApJ...679...17C,2010ApJ...712.1378P}.

The trend is clear: researchers want to find many more lenses and
model them in more sophisticated ways.

@@

\begin{figure}
  \centering
  \includegraphics[width=\myplotswidth]{fig/006941_input}
  \includegraphics[width=\myplotswidth]{fig/006941_arr_time} \\
  \vskip-20pt
  \includegraphics[width=\myplotswidth]{fig/ASW000102p_006941_arriv} 
  \includegraphics[width=\myplotswidth]{fig/006941_spaghetti} \\
  \includegraphics[width=\myplotswidth]{fig/ASW000102p_006941_kappa}
  \includegraphics[width=\myplotswidth]{fig/006941_mass}
  \caption[result 6941 (ASW000102p)]{A simulated lens that mimics a
    lensed quasar, and model results.  The left panels derive from the
    simulation, and the right panels are \spl output.  Details of
    individual panels are in \ref{sec:example_models}.}

>>>>

The first discoveries of gravitational lenses, in each case a galaxy
lensing a background quasar into multiple images
\citep{1979Natur.279..381W,1980Natur.285..641W}, brought in their wake
the first work on lens modelling
\citep{1981ApJ...244..723Y,1981ApJ...244..736Y}.  For one of the
lenses, mass modelling predicted that one of the lensed quasar images
seen would split further into two at higher resolution, which was
crucial in subsequently confirming that the object is indeed an
example of lensing.  It had long been expected that galaxies must
sometimes cause lensed images \citep{1937ApJ....86..217Z}, and it even
been argued that the phenomenon could help measure cosmological
parameters \citep{1964MNRAS.128..307R,1966MNRAS.132..101R}.  But
apparently nobody was expecting that lenses would need detailed
modelling.  In hindsight, the reason is not difficult to understand.
The scale of image separations is set by the angular Einstein radius
\begin{equation}
\sim \left(\frac{4GM}{c^2 D_L}\right)^{1/2}
\simeq 0.1'' \left(\frac{M}{M_\odot}\right)^{1/2}
             \left(\frac{D_L}{\rm pc}\right)^{-1/2}
\end{equation}
where $D_L$ is the distance to the lens, and $M$ its mass.  For
stellar-mass lenses, the image separation is much larger than the size
of the lensing mass, so lenses are well approximated by point masses.
In galaxy lenses, the image separation is comparable to the size of
the galaxy; typically the lensed images are seen through the galaxy
halo.  Consequently, the lensed images depend on the detailed mass
distribution of the lensing galaxy.  Once observations of the lensed
images became available, the need to model the lensing-mass
distribution in detail became obvious.

Nowadays, detailed modelling of the mass distribution is part of any
research using multiply-imaging lenses.  Many works combine multiple
lenses with different mass ranges and other properties.  The largest
single study so far used 58 lenses \citep{2009ApJ...703L..51K} to
uncover the distribution of dark matter around galaxies.
\cite{2011ApJ...740...97L} combined lens models of 21 galaxies with
models of their stellar populations, to study the relation between
stars and dark matter in galaxies.  \cite{2014MNRAS.437..600S}
modelled 18 time-delay lenses together to infer cosmological
parameters.  But lens modelling has remained a separate process from
lens discovery.

@@

\begin{figure}
  \centering
  \includegraphics[width=\myplotswidth]{fig/006941_input}
  \includegraphics[width=\myplotswidth]{fig/006941_arr_time} \\
  \includegraphics[width=\myplotswidth]{fig/ASW000102p_006941_arriv} 
  \includegraphics[width=\myplotswidth]{fig/006941_spaghetti} \\
  \includegraphics[width=\myplotswidth]{fig/ASW000102p_006941_kappa}
  \includegraphics[width=\myplotswidth]{fig/006941_mass}
  \caption[result 6941 (ASW000102p)]{A simulated lens that mimics a
    lensed quasar, and model results.  The left panels derive from the
    simulation, and the right panels are \spl output.  Details of
    individual panels are in \ref{sec:example_models}.}
